### 2020-06-28 neutronics.tex
Transients and conclusion chapter. Update bib. Add figures. Minor edits.

We calculated estimates of $k_{\text{eff}}$ values in Serpent and Moltres for
the static \gls{MSFR} model. Moltres solves the six-group neutron diffusion equations as a
steady-state eigenvalue problem to find the $k_{\text{eff}}$ as it does not
currently have the capability to calculate the adjoint flux. Table
\ref{table:keff} shows the $k_{\text{eff}}$ values from Serpent and Moltres at
973 K and the corresponding salt density, while Table \ref{table:ktemp} shows
the $k_{\text{eff}}$ values for other temperatures at 100 K intervals. We
observe small discrepancies on the order of 100 pcm between the two codes
which we attribute to two main factors: the accuracy of the neutron diffusion
model, and the omission of the blanket tank structural material. The neutron
diffusion model is not as accurate as the other S$_{\text{N}}$ or
SP$_{\text{N}}$ deterministic methods nor the Monte Carlo approach in Serpent.
Regarding the omission of the blanket tank material, we replaced the
2 cm-thick structural material with blanket salt. This replacement may be
partly responsible for the higher $k_{\text{eff}}$ value calculated by Moltres
as the macroscopic fission cross sections of the blanket salt are non-zero for
the higher neutron energy groups. Nevertheless, the discrepancy is smaller
than the 228.5 pcm and 256.7 pcm discrepancies reported by Cervi et al.
\cite{cervi_development_2019} for their six-group $SP_3$ and neutron
diffusion methods, respectively, in OpenFOAM. Aufiero et al.
\cite{aufiero_development_2014} used the same neutron diffusion model in
OpenFOAM for their transient analysis of the \gls{MSFR}, albeit with one
neutron energy group to reduce computational load.

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=.8\textwidth]{reactivity}
    \caption{Reactivity values from Serpent and Moltres. The Doppler
    reactivity values were calculated at a fixed density of 4.1249 g
    cm$^{-3}$. The density reactivity values were calculated at a fixed
    temperature of 973 K.}
    \label{fig:reactivity}
\end{figure}
%
\begin{table}[htb!]
	\centering
	\caption{Doppler, density, and total temperature coefficients
	for the temperature range of 800 K to 1400 K.}
	\begin{tabular}{l S S S S}
		\toprule
		{Code} & {$\alpha_D$ (log) [pcm]} & {$\alpha_D$ (linear) [pcm
		K$^{-1}$]} & {$\alpha_\rho$ [pcm K$^{-1}$]} & {$\alpha_T$ [pcm
		K$^{-1}$]} \\
		\midrule
		{Serpent} & -4034 \pm 14 & 3.737 \pm 0.013 & 3.424 \pm 0.013 & 7.165
		\pm 0.013 \\
		{Moltres} & {-} & {-} & {-} & 7.184\\
		\bottomrule
	\end{tabular}
	\label{table:alpha}
\end{table}

@@

In summary, Moltres replicated most of the relevant neutronics parameters
accurately with the group constant data from Serpent. The $\beta_{\text{eff}}$
and temperature reactivity coefficients, which are important
parameters for modeling transient reactor behavior, are in excellent
agreement. While the
$k_{\text{eff}}$ values have discrepancies on the order of 100 pcm, they are
relatively small compared to the $k_{\text{eff}}$ values from the neutron
diffusion and SP3 models in OpenFOAM \cite{cervi_development_2019}.

>>>>

We calculated estimates of $k_{\text{eff}}$ values in Serpent and Moltres for
the static \gls{MSFR} model. Moltres solves the six-group neutron diffusion equations as a
steady-state eigenvalue problem to find the $k_{\text{eff}}$ as it does not
currently have the capability to calculate the adjoint flux. Table
\ref{table:keff} shows the $k_{\text{eff}}$ values from Serpent and Moltres at
973 K and the corresponding salt density, and Table \ref{table:ktemp} shows
the $k_{\text{eff}}$ values for other temperatures at 100 K intervals. We
observe small discrepancies on the order of 100 pcm between the two codes
which we attribute to two main factors: the accuracy of the neutron diffusion
model, and the omission of the blanket tank structural material. The neutron
diffusion model is not as accurate as the other S$_{\text{N}}$ or
SP$_{\text{N}}$ deterministic methods nor the Monte Carlo approach in Serpent.
Regarding the omission of the blanket tank material, we replaced the
2 cm-thick structural material with blanket salt. This replacement may be
partly responsible for the higher $k_{\text{eff}}$ value calculated by Moltres
as the macroscopic fission cross sections of the blanket salt are non-zero for
the higher neutron energy groups. Nevertheless, the discrepancy is smaller
than the 228.5 pcm and 256.7 pcm discrepancies reported by Cervi et al.
\cite{cervi_development_2019} for their six-group $SP_3$ and neutron
diffusion methods, respectively, in OpenFOAM. Aufiero et al.
\cite{aufiero_development_2014} used the same neutron diffusion model in
OpenFOAM for their transient analysis of the \gls{MSFR}, albeit with one
neutron energy group to reduce computational load.

@@

\begin{figure}[htb!]
    \centering
    \includegraphics[width=.8\textwidth]{reactivity}
    \caption{Reactivity values from Serpent and Moltres. The Doppler
    reactivity values were calculated at a fixed density of 4.1249 g
    cm$^{-3}$. The density reactivity values were calculated at a fixed
    temperature of 973 K.}
    \label{fig:reactivity}
\end{figure}
%
\begin{table}[htb!]
	\centering
	\caption{Doppler, density, and total temperature coefficients
	for the temperature range of 800 K to 1400 K.}
	\begin{tabular}{l S S S S}
		\toprule
		{Code} & {$\alpha_D$ (log) [pcm]} & {$\alpha_D$ (linear) [pcm
		K$^{-1}$]} & {$\alpha_\rho$ [pcm K$^{-1}$]} & {$\alpha_T$ [pcm
		K$^{-1}$]} \\
		\midrule
		{Serpent} & -4034 \pm 14 & -3.737 \pm 0.013 & -3.424 \pm 0.013 &
		-7.165 \pm 0.013 \\
		{Moltres} & {-} & {-} & {-} & -7.184\\
		\bottomrule
	\end{tabular}
	\label{table:alpha}
\end{table}

@@

In summary, Moltres replicated most of the relevant neutronics parameters
accurately with the group constant data from Serpent. The $\beta_{\text{eff}}$
and temperature reactivity coefficients, which are important
parameters for modeling transient reactor behavior, are in excellent
agreement. The $k_{\text{eff}}$ values have discrepancies on the order of 100
pcm and they are relatively small compared to the $k_{\text{eff}}$ values from
the neutron diffusion and SP3 models in OpenFOAM
\cite{cervi_development_2019}.

### 2020-06-28 steadystate.tex
Transients and conclusion chapter. Update bib. Add figures. Minor edits.

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{.365\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.625\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp-fiorina}
    \end{subfigure}
    \caption{Temperature and velocity fields in the core from Moltres
    (left), Polimi (center), and TUDelft (right) models. The colors represent
    temperature according to the respective colorbars while the arrows
    represent velocity fields.}
    \label{fig:flow-temp}
\end{figure}

@@

While the temperatures at the hotspots are well below the melting point of the
Ni-alloy structure (1500 K), they may cause undue thermal stress on the
blanket tank structure and induce relatively faster salt corrosion rates. A
sudden, large reactivity insertion could push fuel salt temperatures above the
melting point of the Ni-alloy and cause irreversible damage. Furthermore,
the reservoir of hot fuel salt may cause unpredictable behavior during
transient scenarios when the flow profile undergoes a drastic change.
Thus, Rouch et al. \cite{rouch_preliminary_2014} developed an improved
hourglass-shaped design to optimize flow distribution and prevent these
recirculation zones and hotspots from forming. A study of this new design
using Moltres is a potential subject for future work when a proper turbulence
model is in place.

@@

\begin{figure}[b!]
    \centering
    \begin{subfigure}[t]{.30\textwidth}
        \centering
        \vspace{.9cm}
        \includegraphics[width=\textwidth]{ss-pre}
    \end{subfigure}
    \begin{subfigure}[t]{.69\textwidth}
        \centering
        \vspace{0pt}
        \includegraphics[width=\textwidth]{fiorina-pre}
    \end{subfigure}
    \caption{Total \gls{DNP} distribution in the core from Moltres
    (left), Polimi (center), and TUDelft (right) models.}
    \label{fig:pre}
\end{figure}

@@

The fraction of out-of-core emissions from our Moltres model
differs significantly by approximately 10\%, while $\beta_c$ differs by
60-70 pcm. We attribute the former to the lesser \gls{DNP} retention in the
stagnant flow regions in the core; the \glspl{DNP} are more evenly distributed
along the entire primary loop, leading to more delayed neutron emissions in
the outer loop region. The most likely reason for this is differences in the
flow pattern in the recirculation zone because convective species transport
dominates diffusive effects in the \gls{MSFR}. The exact flow pattern in the
recirculation zones in the Polimi and TUDelft models is likely to differ from
that in Moltres (Figure \ref{fig:flow}). Figure \ref{fig:flow-temp} also shows
some minor differences in the magnitude of the flow in the recirculation zones
between Moltres, and the Polimi and TUDelft models. Although the sizes of the
arrows representing flow velocity are not normalized to the same scales, a
quick comparison between the largest arrows and the arrows in the
recirculation zone indicate that the recirculation zone in Moltres is
relatively more stagnant. This could explain the concentration of \glspl{DNP}
along an ``arc'' closer to the center of the core in Moltres as opposed to the
more even distribution of \glspl{DNP} throughout the whole recirculation zones
in the Polimi and TUDelft models. The higher peak \glspl{DNP} distribution in
Moltres also supports this assertion.

In spite of the greater delayed neutron losses, the $\beta_c$ value is higher
in Moltres than the Polimi and TUDelft models. To account for this
peculiarity, we note that Fiorina et al. \cite{fiorina_modelling_2014}
applied adjoint flux weighting for their $\beta_c$ calculation while we report
our value as the physical fraction without adjoint weighting. The weighting
results in a significant difference in $\beta_c$ because a large fraction of
the \glspl{DNP} decay in the recirculation zones, where the neutron importance
is noticeably diminished.

>>>>

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{.365\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.625\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp-fiorina}
    \end{subfigure}
    \caption{Temperature and velocity fields in the core from Moltres
    (left), Polimi (center), and TUDelft (right) models. The colors represent
    temperature according to the respective colorbars and the arrows
    represent velocity fields.}
    \label{fig:flow-temp}
\end{figure}

@@

Although the temperatures at the hotspots are well below the melting point of the Ni-alloy structure (1500 K), they may cause undue thermal stress on the
blanket tank structure and induce relatively faster salt corrosion rates. A
sudden, large reactivity insertion could push fuel salt temperatures above the
melting point of the Ni-alloy and cause irreversible damage. Furthermore,
the reservoir of hot fuel salt may cause unpredictable behavior during
transient scenarios when the flow profile undergoes a drastic change.
Thus, Rouch et al. \cite{rouch_preliminary_2014} developed an improved
hourglass-shaped design to optimize flow distribution and prevent these
recirculation zones and hotspots from forming. A study of this new design
using Moltres is a potential subject for future work when a proper turbulence
model is in place.

@@

\begin{figure}[b!]
    \centering
    \begin{subfigure}[t]{.30\textwidth}
        \centering
        \vspace{.9cm}
        \includegraphics[width=\textwidth]{ss-pre}
    \end{subfigure}
    \begin{subfigure}[t]{.69\textwidth}
        \centering
        \vspace{0pt}
        \includegraphics[width=\textwidth]{fiorina-pre}
    \end{subfigure}
    \caption{Total delayed neutron source distribution in the core from
    Moltres (left), Polimi (center), and TUDelft (right) models.}
    \label{fig:pre}
\end{figure}

@@

The fraction of out-of-core emissions from our Moltres model
differs significantly by approximately 10\%, and $\beta_c$ differs by
60-70 pcm. We attribute the former to the lesser \gls{DNP} retention in the
stagnant flow regions in the core; the \glspl{DNP} are more evenly distributed
along the entire primary loop, leading to more delayed neutron emissions in
the outer loop region. The most likely reason for this is differences in the
flow pattern in the recirculation zone because convective species transport
dominates diffusive effects in the \gls{MSFR}. The exact flow pattern in the
recirculation zones in the Polimi and TUDelft models is likely to differ from
that in Moltres (Figure \ref{fig:flow}). Figure \ref{fig:flow-temp} also shows
some minor differences in the magnitude of the flow in the recirculation zones
between Moltres, and the Polimi and TUDelft models. Although the sizes of the
arrows representing flow velocity are not normalized to the same scales, a
quick comparison between the largest arrows and the arrows in the
recirculation zone indicate that the recirculation zone in Moltres is
relatively more stagnant. This could explain the concentration of \glspl{DNP}
along an ``arc'' closer to the center of the core in Moltres as opposed to the
more even distribution of \glspl{DNP} throughout the whole recirculation zones
in the Polimi and TUDelft models. The higher peak \glspl{DNP} distribution in
Moltres also supports this assertion.

In spite of the greater delayed neutron losses, the $\beta_c$ value is higher
in Moltres than the Polimi and TUDelft models. To account for this
peculiarity, we note that Fiorina et al. \cite{fiorina_modelling_2014}
applied adjoint flux weighting for their $\beta_c$ calculation whereas we
report our value as the physical fraction without adjoint weighting. The
weighting results in a significant difference in $\beta_c$ because a large
fraction of the \glspl{DNP} decay in the recirculation zones, where the
neutron importance is noticeably diminished.

### 2020-06-28 transient.tex
Transients and conclusion chapter. Update bib. Add figures. Minor edits.

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.85\textwidth]{lohs-heat}
    \caption{Power output during
    an unprotected loss of heat sink transient in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:lohsheat}
    \includegraphics[width=.85\textwidth]{lohs-temp}
    \caption{Average core temperature increase during
    an unprotected loss of heat sink transient in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:lohstemp}
\end{figure}

@@

A loss of forced flow transient can occur in the event of a station
blackout; the pumps would cease operating due to the loss of AC electrical
power. Natural circulation resulting from temperature-dependent density
changes becomes the dominant driving force for salt flow in the primary loop.
While Fiorina et al. \cite{fiorina_modelling_2014} applied the Boussinesq
approximation for bouyancy-driven flow in their models, it is not directly
applicable to the \gls{MSFR} model in this thesis because we partitioned the
primary loop into two separate geometries and used Dirichlet boundary
conditions at the inlet to drive flow. Fiorina et al.'s Polimi and TUDelft
models featured complete exponential coast-downs of the pumps with a time
constant of 5 s. The resulting flow rate from natural circulation was
approximately 18 times smaller than the initial flow rate. Figure
\ref{fig:flowrate} shows that the actual flow rate decreased with a time
constant of 8 s. Thus, for the \gls{MSFR} model in Moltres, we imposed a
similar exponential decay term with a time constant of 8 s on the inflow
Dirichlet boundary condition:
%
\begin{align}
    \text{Flow rate, } v = 0.25862 + (4.5-0.25862) e^{-t/8} 
    \label{eq:flowrate}
\end{align}

@@

The reduced flow rate also decreases the heat transfer rate between the
primary and intermediate loop through the heat exchanger as the heat transfer
coefficient is dependent on the flow rate. This step was problematic because
the pointwise heat exchanger implementation in Moltres performs differently
compared to the heat exchangers that take up ``36\% of the out-of-core part''
in the Polimi and TUDelft models \cite{fiorina_modelling_2014}. Most of the
cooling happens in the top half of the heat exchanger where the temperature
differential between the primary and intermediate loops is the largest. In the
Polimi and TUDelft models, the overall heat transfer coefficient is the
``harmonic mean of the heat transfer coefficients on each side of the heat
exchanger''. For this loss of flow transient, the authors intended to focus on
the primary loop and assumed that only the pumps in the primary loop failed.
In addition to this, the authors applied the Dittus-Boelter correlation (cite
*) for
the relationship between the primary side heat transfer coefficient and the
flow rate. The Dittus-Boelter correlation for fluids being cooled is:
%
\begin{align}
    Nu &= 0.023 Re^{0.8} Pr^{0.3} \label{eq:db} \\
    \intertext{where}
    Nu &= \text{ Nusselt number,} \nonumber \\
    Re &= \text{ Reynolds number,} \nonumber \\
    Pr &= \text{ Prandtl number.} \nonumber
\end{align}

The only direct relation to flow velocity in the Dittus-Boelter correlation is
through the $Re$ term, which is directly proportional to flow velocity. This
gives the following relation between the heat transfer coefficient $h$ and
flow velocity $v$:
%
\begin{align}
    h \propto v^{0.8} \label{eq:hv}
\end{align}
%
However, this relation provided very different results in the unprotected loss
of flow and pump overspeed transients compared to the Polimi and TUDelft
models; it overpredicted the equilibrium power output in the
unprotected loss of flow transient and underpredicted the same parameter in
the unprotected pump overspeed transient. Upon further investigation, we found
that raising the power of $v$ from 0.8 to 1.1 brought the average core
temperatures closer to the results from the other models in both transients.
Therefore, we adopted the raised power in this study.

Another issue pertained to the turbulent viscosity $\mu_t$ as a function of
$v$. When we made a simple approximation of $\mu_t$ being directly
proportional to $v$, the results
differed significantly compared to the Polimi and TUDelft models (Appendix *).
This may have been due to buoyancy-driven flow contributing to turbulence; the
turbulent energy $k$ equation in the $k$-$\epsilon$ model has an explicit
source term from buoyancy effects (Cite *). Another point to note is the
Reynolds number does not change if $\mu$ and $v$ decrease in tandem. This
preserves the existence of the recirculation zone in the core and it is at
odds with the results from the Polimi and TUDelft models, which show that the
recirculation zones disappear during the loss of flow transient. We
circumvented this issue by letting higher fractions of $\mu_t$ be conserved
regardless of the final flow velocity. This measure allowed for laminar flow
to develop in the core and yielded results resembling those from the Polimi
and TUDelft models.

Figures * show

@@

In the context of this thesis, pump overspeed refers to a sustained
increase in pump speed in the primary coolant loop. The increased flow rate
impacts reactor performance in several ways.
It affects the neutronics by reducing the in-core $\beta_eff$ as more of the
shorter-lived precursors will tend to flow out of the core before decaying.
This net loss of neutrons reduces the reactivity in the core, thereby causing
core temperatures to fall to counteract this change through temperature
reactivity feedback. The increased flow rate also enhances the heat transfer
coefficient on the primary loop side of the heat exchanger and enables the
reactor to operate at a higher power output. At the same time, the improved
mixing flattens the temperature distribution in the core.

We followed Fiorina et al.'s implementation \cite{fiorina_modelling_2014} by
ramping up the inlet velocity, $u$, by 50\% from the nominal value, $u_0$,
according to the following formula:

\begin{align}
    u(t) &= u_0 [1 + 0.5 (1 - e^{-t / \tau})] \\
    \intertext{where}
    \tau &= 5 \text{ s.} \nonumber
\end{align}
%
For this transient, we assumed that $\mu_t$ was directly proportional to $v$
because the buoyancy effects are negligible and the recirculation zones 
persist throughout the entire duration. Figures \ref{fig:poheat} and
\ref{fig:potemp} show the power output and average core temperature increase
during the unprotected pump overspeed transient in the Moltres, Polimi, and
TUDelft models.

Unlike in the loss of flow transient, the $\mu_t \propto v$ approximation
worked well in pump overspeed transient.

@@

>>>>

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.85\textwidth]{lohs-heat}
    \caption{Power output during
    an unprotected loss of heat sink transient in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014} without decay heat.}
    \label{fig:lohsheat}
    \includegraphics[width=.85\textwidth]{lohs-temp}
    \caption{Average core temperature increase during
    an unprotected loss of heat sink transient in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014} without decay heat.}
    \label{fig:lohstemp}
\end{figure}

@@

A loss of forced flow transient can occur in the event of a station
blackout; the pumps would cease operating due to the loss of AC electrical
power. Natural circulation resulting from temperature-dependent density
changes becomes the dominant driving force for salt flow in the primary loop.
Fiorina et al. \cite{fiorina_modelling_2014} applied the Boussinesq
approximation for bouyancy-driven flow in their models, but we were unable to
follow this procedure because we partitioned the
primary loop into two separate geometries and used Dirichlet boundary
conditions at the inlet to drive flow. Fiorina et al.'s Polimi and TUDelft
models featured complete exponential coast-downs of the pumps with a time
constant of 5 s. The resulting flow rate $\dot{m}$ from natural circulation was
approximately 18 times smaller than the initial $\dot{m}$. Figure
\ref{fig:flowrate} shows that the actual $\dot{m}$ decreased with a time
constant of 8 s. Thus, for the \gls{MSFR} model in Moltres, we imposed a
similar exponential decay term with a time constant of 8 s on the inflow
Dirichlet boundary condition:
%
\begin{align}
    \text{Flow rate, } v = 0.25862 + (4.5-0.25862) e^{-t/8} 
    \label{eq:flowrate}
\end{align}

@@

The reduced $\dot{m}$ also decreases the heat transfer rate between the
primary and intermediate loop through the heat exchanger as the heat transfer
coefficient $h$ is dependent on the $\dot{m}$. This step was problematic
because
the pointwise heat exchanger implementation in Moltres performs differently
compared to the heat exchangers that take up ``36\% of the out-of-core part''
in the Polimi and TUDelft models \cite{fiorina_modelling_2014}. Most of the
cooling happens in the top half of the heat exchanger where the temperature
differential between the primary and intermediate loops is the largest. In the
Polimi and TUDelft models, the overall $h$ is the
``harmonic mean of the heat transfer coefficients on each side of the heat
exchanger''. For this loss of flow transient, the authors intended to focus on
the primary loop and assumed that only the pumps in the primary loop failed.
In addition to this, the authors applied the Dittus-Boelter correlation
\cite{dittus_heat_1930} for the relationship between the primary side $h$ and
$\dot{m}$. The Dittus-Boelter correlation for fluids being cooled is:
%
\begin{align}
    Nu &= 0.023 Re^{0.8} Pr^{0.3} \label{eq:db} \\
    \intertext{where}
    Nu &= \text{ Nusselt number,} \nonumber \\
    Re &= \text{ Reynolds number,} \nonumber \\
    Pr &= \text{ Prandtl number.} \nonumber
\end{align}
%
The only direct relation to $\dot{m}$ in the Dittus-Boelter correlation is
through the $Re$ term, which is directly proportional to flow velocity $v$.
This gives the following relation between $h$ and$v$:
%
\begin{align}
    h \propto v^{0.8} \label{eq:hv}
\end{align}
%
However, this relation provided very different results in the unprotected loss
of flow and pump overspeed transients compared to the Polimi and TUDelft
models; it overpredicted the equilibrium power output in the
unprotected loss of flow transient and underpredicted the same parameter in
the unprotected pump overspeed transient. Upon further investigation, we found
that raising the power of $v$ from 0.8 to 1.1 brought the average core
temperatures closer to the results from the other models in both transients.
Therefore, we adopted the raised power in this study.

Another issue pertained to the turbulent viscosity $\mu_t$ as a function of
$v$. When we made a simple approximation of $\mu_t$ being directly
proportional to $v$, the results differed significantly compared to the Polimi
and TUDelft models (Appendix *). This is likely due to buoyancy-driven flow
contributing to turbulence; the turbulent energy $k$ equation in COMSOL's
$k$-$\epsilon$ model has an explicit source term from buoyancy effects
\cite{comsol_ab_comsol_2018}. Another point to note is the
Reynolds number does not change if $\mu$ and $v$ decrease in tandem. This
preserves the existence of the recirculation zone in the core and it is at
odds with the results from the Polimi and TUDelft models, which show that the
recirculation zones disappear during the loss of flow transient. We
circumvented this issue by letting fixed fractions of the initial $\mu_{t,0}$
be conserved regardless of the final flow velocity, according to the following
equation:
%
\begin{align}
    \mu_t &= \mu_c + (\mu_{t,0} - \mu_c) e^{-t/8} \label{lofmu} \\
    \intertext{where}
    \mu_c &= \text{ conserved fraction of $\mu_{t,0}$ [Pa$\cdot$s].} \nonumber
\end{align}
%
This measure allowed for laminar flow to develop in the core and yielded
results with closer resemblance to those from the Polimi and TUDelft models.

Figures \ref{fig:lofheat} and \ref{fig:loftemp} show the power output and
average core temperature increase during the unprotected loss of flow
transient in the Moltres, Polimi, and TUDelft models without decay heat
modeling. The three sets of results from Moltres correspond to $\mu_c =
\frac{1}{4} \mu_{t,0}, \frac{1}{2} \mu_{t,0}, \text{and } \frac{3}{4}
\mu_{t,0}$. Moltres
performs poorer in this transient relative to the two previous transients.
Although Moltres shows the same decreasing trend in power output, it failed to
capture the exact individual features in the reactor response. In the Polimi 
and TUDelft models, Fiorina et al.
stated that after around $t=15$ s, the ``flow pattern changed in the core and
the recirculation zones started to disappear''. As a result, there is a sudden
drop in the average core temperature as the pocket of hot salt leaves the
core. In Moltres, the wider peak in the average core temperature
indicates that there was a more gradual change in the flow pattern.

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.85\textwidth]{lof-heat}
    \caption{Power output during
    an unprotected loss of flow transient in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:lofheat}
    \includegraphics[width=.85\textwidth]{lof-temp}
    \caption{Average core temperature increase during
    an unprotected loss of flow transient in the Moltres, Polimi, and
    TUDelft models \cite{fiorina_modelling_2014}.}
    \label{fig:loftemp}
\end{figure}

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.37\textwidth]{lof-flow-temp}
    \includegraphics[width=.62\textwidth]{fiorina-lof}
    \caption{Temperature and velocity fields in the core at $t=300$ s during
    a loss of flow transient in the Moltres ($\mu_c = \frac{1}{2} \mu_{t,0}$),
    Polimi, and TUDelft models.}
    \label{fig:lofflowtemp}
\end{figure}

Figure \ref{fig:lofflowtemp} shows the flow patterns and temperature
distribution in the core at $t=300$ s in all three models. Figures
\ref{fig:lofheat}, \ref{fig:loftemp}, and \ref{fig:lofflowtemp} combined
highlight the difference between laminar flow in the Moltres model and
buoyancy-driven flow in the other models, and its impact on the reactor
response. They show that low-speed laminar flow is a poor substitute for
buoyancy-driven flow in the context of the MSFR. It is particularly evident
in the transition from high-speed turbulent flow to low-speed viscous
flow as Moltres did not capture the intermediate stages accurately. The
simplifying assumption for the uniform, time-dependent $\mu_t$ is also not
ideal for a safety analysis code.

The results from this transient inform our goals for Moltres: 1)
implementing a proper turbulence model, and
2) developing a new heat exchanger feature that is compatible with the
buoyancy-driven flow capabilities already present in Moltres.

@@

Pump overspeed refers to a sustained
increase in pump speed in the primary coolant loop. The increased flow rate
$\dot{m}$ impacts reactor performance in several ways.
It affects the neutronics by reducing the in-core $\beta$ as more of the
shorter-lived precursors will tend to flow out of the core before decaying.
This net loss of neutrons reduces the reactivity in the core, thereby causing
core temperatures to fall to counteract this change through temperature
reactivity feedback. The increased $\dot{m}$ also enhances the heat transfer
coefficient on the primary loop side of the heat exchanger and enables the
reactor to operate at a higher power output. At the same time, the improved
mixing flattens the temperature distribution in the core.

We followed Fiorina et al.'s implementation \cite{fiorina_modelling_2014} by
ramping up the inlet velocity, $u$, by 50\% from the nominal value, $u_0$,
according to the following formula:
%
\begin{align}
    u(t) &= u_0 [1 + 0.5 (1 - e^{-t / \tau})] \label{eq:overspeed}
    \intertext{where}
    \tau &= 5 \text{ s.} \nonumber
\end{align}
%
For this transient, we assumed that $\mu_t$ was directly proportional to $v$
because the buoyancy effects are negligible and the recirculation zones 
persist throughout the entire duration. Figures \ref{fig:poheat} and
\ref{fig:potemp} show the power output and average core temperature increase
during the unprotected pump overspeed transient in the Moltres, Polimi, and
TUDelft models. Figure \ref{fig:poshort} shows the same results for the first
20 seconds of the transient. At the start of the transient, the rising flow
rate cools the core and causes power output to rise sharply. Although the
average core temperature has a strictly decreasing trend, the temperature at
the center of the core briefly rises due ot the sharp increase in power
output. Since this is the region where most of the fissions take place, the
Doppler effect and salt expansion causes the power output to stall and dip
briefly before starting to rise again at $t=2.5$ s. The reactor tends to a new
equilibrium power output and average core temperature. The temperature
distribution in the core is more evenly distributed because the turbulent
thermal conductivity $k_t$ is directly proportional to $mu_t$.

@@

\begin{figure}[htbp!]
    \centering
    \begin{subfigure}[t]{.485\textwidth}
        \centering
        \includegraphics[width=\textwidth]{po-heat-short}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.485\textwidth}
        \centering
        \includegraphics[width=\textwidth]{po-temp-short}
    \end{subfigure}
    \caption{The first 20 s of the power output and average core temperature
    increase during an unprotected pump overspeed transient.}
    \label{fig:poshort}
\end{figure}

The results show excellent agreement with the Polimi and TUDelft models. The
average core temperature increase in particular reproduces the Polimi and   
TUDelft results very well and the curve falls between the other two curves.
The power output is higher because the Moltres \gls{MSFR} model has a
stronger negative temperature coefficient than the other two models.
